### 2020-01-22 prioritizing.tex
label the section

\section{Prioritizing Mutants}
\label{sec:prioritizing}